Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-10 paper/paper.tex
Updated

Let   us  consider   determining  the   liveness  of   a   closure  at
runtime. Consider  a closure  $(app~f~y)$ has a  reference from
the root  set variable $x$. If $x$  is live,  then we  know that the
cell $(app~f~y)$ has  to be copied. To find out  the
liveness of $y$, we need the program point which triggered the
evaluation of $(app~f~y)$ and pick up the liveness of $y$ from this
program point. However, this is only possible if the activation record
of the function which created this closure is still on the activation
stack. This does not always happen since a function could pack a
closure in a \CONS\ cell and return the \CONS\ cell.

An  alternative  could  be   to  carry  extra  information  with  each
closure---the {\em free variables in the closure and the program point
  at  which  it  was  evaluated}.   This  gives  us  precise  liveness
information  of the  free  variables  of a  closure  but involves  the
overhead of  updating the program point  information during execution.
To avoid this  overhead we go for a safe  approximation by storing the
{\em creation  point of the closure}  and using it  to check liveness.
This is safe  since the liveness at the creation  point of the closure
dominates all possible evaluation points  of the closure, but it might
be imprecise  as the actual  liveness might be  way less than  what is
found at the creation point.

>>>>

Let   us  consider   determining  the   liveness  of   a  closure   at
runtime. Consider a closure $(app~f~y)$  has a reference from the root
set  variable  $x$.  If $x$  is  live,  then  we  know that  the  cell
$(app~f~y)$ has  to be copied.  To copy  the heap structure  rooted at
$y$, we need the liveness of  $y$ at the program point which triggered
the evaluation of  $(app~f~y)$. However, this is only  possible if the
activation record of the function  which created this closure is still
on the activation stack. This does  not always happen since a function
could pack a closure in a \CONS\ cell and return the \CONS\ cell.

An alternative could be to  carry with each closure, extra information
about the program point which triggered its evaluation.  This gives us
precise liveness  information of the  free variables of a  closure but
involves the overhead of updating the program point information during
execution.  To avoid  this overhead we go for a  safe approximation by
storing the {\em creation point of  the closure} and using it to check
liveness.  This  is safe since the  liveness at the creation  point of
the closure dominates  all possible evaluation points  of the closure,
but it  might be imprecise  as the actual  liveness might be  way less
than what is found at the creation point.
